---
title: "Transmission model simulation"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    source_code: embed
    theme: united
runtime: shiny
---

```{r global, include=FALSE, warning=F, message=F}
# load data in 'global' chunk so it can be shared by all users of the dashboard
library(tidyverse)
library(DiagrammeR)
library(plotly)
library(deSolve)
library(shinyWidgets)
colorsset <- c("lightsteelblue", "indianred4", "slateblue1")
```

## Inputs {.sidebar}

This dashboard simulates the progression of a disease in a population using a **S**usceptible-**I**nfected-**R**ecovered-**S**usceptible deterministic ordinary differential equation model. Adjust the parameters below based on literature, prior knowledge, or data.


```{r}
sliderInput("input_prev", "Initial prevalence:", value = .01, min = .01, max = 1, step = .1)

sliderInput("input_trans", "Transmission probability:", value = .01, min = 0, max = 1, step = .1)

numericInput("input_reco", "Infection duration (weeks):", value = 1, min = 0)

numericInput("input_wane", "Waning immunity (weeks immune):", value = 2, min = 0)

pickerInput("vax_age",
            "Age group(s) to vaccinate:",
            choices =  c("None"=1, "Youth (0-18)"=2, "Adults (19-64)"=3, "Older adults (65+)"=4),
            multiple = TRUE,
            selected = 1,
            options = list(
                  title = "Select variables",
                  `actions-box` = TRUE,
                  `deselect-all-text` = "Remove"
                ))

sliderInput("input_vax", "Vaccine uptake (proportion of susceptibles vaccinated each week):", value = 0, min = 0, max = 1, step = .1)

checkboxInput("ui", "Add uncertainty by using ranges of values for transmission probability instead of a fixed value", value = FALSE)
```


Column
-------------------------------------
### Compartments over time
```{r}
noiz <- .1
    
reactiveData <- reactive({

  # Define the parameters !!To have the timestep be a day, will multiply the week values by 7
  prev <- input$input_prev
  trans <- input$input_trans
  reco <-  1/(input$input_reco*7)
  wane <- 1/(input$input_wane*7)
  vax <- input$input_vax/7 # divide since this is proportion per week
  age <- input$vax_age
  ui <- input$ui
  
  # Setup the proportion vaccinated by age group
  age_dist_df <- data.frame(grp= 1:4, prop = c(0, .221, .614, .165)) # distribution of population taken from Census 2022
  vax_prop <- sum(age_dist_df[unlist(age), "prop"])
  
  # To simulate targeted vaccination, the proportion vaccinated (vax) is multiplied by the proportion of the population being targeted (vax_age). 
  vax <- vax*vax_prop

  SIRS <- function(time, current_state, params){
    
    with(as.list(c(current_state, params)),{
      dS <- -(trans*S*I) - vax * S + wane * R
      dI <- (trans*S*I) - reco*I 
      dR <- reco*I + vax * S - wane * R
      
      return(list(c(dS, dI, dR)))
    })
  }
  
  params <- c(trans, reco, vax, wane)
  
  initial_state <- c(S=1-prev, I=prev, R=0)
  
  times <- 0:30 #days
  
  dats <- data.frame(ode(initial_state, times, SIRS, params))

  # Add random noise if uncertainty is desired - will just add noise to the transmission probability since least likely to be known
  if(isTRUE(ui)){
    # Was adding random noise to the parameters and then rerunning but the intervals were crossing over the point estimate and not making sense so simplified
    # datslow = data.frame(ode(initial_state, times, SIRS, parms = c(trans = pmax(trans - noiz,0), 
    #                                                                reco = pmax(reco - noiz,0), 
    #                                                                vax, 
    #                                                                wane = pmax(wane - noiz,0)))) %>%
    #   dplyr::rename_with(~ paste0("low_", .))
    # datshi = data.frame(ode(initial_state, times, SIRS, parms = c(trans = pmin(trans + noiz, 1), 
    #                                                               reco = pmin(reco + noiz, 1), 
    #                                                               vax, 
    #                                                               wane = pmin(wane + noiz, 1)))) %>%
    #   dplyr::rename_with(~ paste0("hi_", .))
    # dats <- cbind(dats, datslow, datshi)
    dats <- dats %>% 
      mutate(across(c(S, I ,R), ~pmax(. - noiz, 0), .names = "low_{.col}"),
             across(c(S, I ,R), ~pmin(. + noiz,1), .names = "hi_{.col}"))
  }
  
  #run again with no vaccination for baseline
  bSIRS <- function(time, current_state, params){
    
    with(as.list(c(current_state, params)),{
      dS <- -(trans*S*I) + wane * R
      dI <- (trans*S*I) - reco*I 
      dR <- reco*I - wane * R
      
      return(list(c(dS, dI, dR)))
    })
  }
  
  params <- c(trans, reco, wane)
  
  baseline <- ode(initial_state, times, bSIRS, params)
  
  dats$base = data.frame(baseline)$I
  dats
})

renderPlot({
      {if(isTRUE(input$ui)){
        ggplot(reactiveData(), aes(x = time)) +
        geom_ribbon(aes(ymin = low_S, ymax = hi_S, fill = "Susceptible"), alpha = .5) +
        geom_ribbon(aes(ymin = low_I, ymax = hi_I, fill = "Infectious"), alpha = .5) +
        geom_ribbon(aes(ymin = low_R, ymax = hi_R, fill = "Recovered/Immune"), alpha = .5)
      } else {
        ggplot(reactiveData(), aes(x = time))
      }} +
      geom_line(aes(y = S, color = "Susceptible"), linewidth = 3) +
      geom_line(aes(y = I, color = "Infectious"), linewidth = 3) +
      geom_line(aes(y = R, color = "Recovered/Immune"), linewidth = 3) +
      xlab("Days") +
      ylab("Proportion") +
      scale_color_manual(name = "", values=c(colorsset)) +
      scale_fill_manual(name = "", values=colorsset) + 
      scale_y_continuous(limits = c(0, 1)) +
      scale_linewidth_manual(name = "") +
      theme(text = element_text(size = 18))
})
```

### New cases without and with vaccination

```{r}
#make vaccination a difference color!!!
renderPlot({
  reactiveData()%>% 
    mutate("With vaccination" = pmax(diff(c(0, I)), 0),
            Baseline = pmax(diff(c(0, base)), 0), 
            index = 1:n()) %>% 
    # {if(isTRUE(input$ui)){
    #   mutate(., wvaxlo = pmax(diff(c(0, low_I)), 0),
    #          wvaxhi = pmax(diff(c(0, hi_I)), 0)) %>%
    #   pivot_longer(c(Baseline, "With vaccination")) %>%
    #   mutate(wvaxlo = case_when(value < wvaxlo ~ value,
    #                             TRUE ~ wvaxlo),
    #          wvaxlo = case_when(name == "Baseline" ~ NA, # for some reason this is only working when done in two separate steps.
    #                             TRUE ~ wvaxlo),
    #          wvaxhi = case_when(value > wvaxhi ~ value,
    #                             TRUE ~ wvaxhi),
    #          wvaxhi = case_when(name == "Baseline" ~ NA,
    #                             TRUE ~ wvaxhi)) %>% 
    #   ggplot(aes(x = time, y = value*100)) +
    #   geom_ribbon(aes(ymin = wvaxlo*100, ymax = wvaxhi*100, fill = factor(name, labels = c("With vaccination", "Baseline"), levels = c("With vaccination", "Baseline"))), alpha = .5)
    # } else {
      pivot_longer( c(Baseline, "With vaccination")) %>%
      ggplot(aes(x = time, y = value*100)) +
    # }} +
    geom_line(aes(color = factor(name, labels = c("With vaccination", "Baseline"), levels = c("With vaccination", "Baseline"))), linewidth = 3) +
    scale_color_manual(name = "", values=c("lightsteelblue", "steelblue"))+
    # scale_fill_manual(name = "", values=c("lightsteelblue", NA))+
    scale_linewidth_manual(name = "")+
    theme(text = element_text(size = 18)) +
    labs(y = "Incidence per 100,000", x = "Days")
})
 
```

Column
-------------------------------------

### Number of infections without and with vaccination

```{r}
library(tidyverse)
# renderPlot({
#   ggplot(reactiveData() %>% mutate(across(-Time, ~cumsum(.))) %>% pivot_longer(-Time), aes(x = Time)) +
#       geom_area(aes(y = value, fill = name), position = "stack") +
#       xlab("Time") +
#       ylab("People") +
#       scale_fill_manual(name = "", values=c(colorsset))+
#       scale_linewidth_manual(name = "") +
#         theme(text = element_text(size = 18))
# })
renderPlot({
  pdat <- reactiveData() %>% 
    summarise("With vaccination" = pmax(diff(c(0, I)), 0),
              Baseline = pmax(diff(c(0, base)), 0), 
              index = 1:n()) %>%
    pivot_longer(-index) %>% 
    group_by(name) %>%
    summarise(across(everything(), ~sum(.)*100)) 
  pdat %>% 
    ggplot(aes(x = name, y = value, fill = name)) + 
    geom_col(alpha = .8) + 
    scale_fill_manual(values = c("steelblue", "lightsteelblue")) + 
    theme(legend.position = "none", 
          text = element_text(size = 18), 
          axis.title.x = element_text(color = colorsset[2], face = "bold")) +
    labs(x = paste0(round(100*abs(diff(pdat$value))/pdat[pdat$name == "Baseline","value"]) ,"% reduction in cases"), y = "Number of infections\nper 100,000")
})
```

### Diagram

```{r}
# Create a new graph
graph <- create_graph()

# Add nodes for compartments
graph <- graph %>%
  add_node("S", label = "Susceptible", node_aes = node_aes(shape = "square", 
                                                           fillcolor = colorsset[3], fontsize = 5, color = "white")) %>%
  add_node("I", label = "Infectious", node_aes = node_aes(shape = "square", 
                                                          fillcolor = colorsset[1], fontsize = 5, color = "white")) %>%
  add_node("R", label = "Recovered", node_aes = node_aes(shape = "square", 
                                                         fillcolor = colorsset[2], fontsize = 5, color = "white")) %>% 
  
  # Add edges for transitions between compartments
  add_edge(from="Susceptible", to="Infectious", edge_aes = edge_aes(fontsize = 3, 
                                                                    label = "Transmission probability")) %>%
  add_edge(from="Infectious", to="Recovered", edge_aes = edge_aes(fontsize = 3, 
                                                                  label = "1/Infection duration")) %>% 
  add_edge(from="Recovered", to="Susceptible", edge_aes = edge_aes(fontsize = 3, 
                                                                   label = "Waning immunity")) %>% 
  add_edge(from="Susceptible", to="Recovered", edge_aes = edge_aes(fontsize = 3, 
                                                                   label = "Vaccine uptake")) 

# Render the graph
render_graph(graph, layout = "fr", width = 350, height = 350)
```

### Basic reproductive number {data-height=125}

<!-- $$ -->
<!-- R_{0} = \ \ \ \ \  \ \ \ \ \ \  \ \ \ \ \ \ \alpha\ \ \ \ \ \ \  \ \ \ \ \ *\ \ \ \ \ \ \  \ \ \ \ \  1/\gamma \  \ \ \ \ \ \ \ \ \ \ * \ \ \ \ \ \  \ \ \ \ \ \beta \\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Contact\ rate \ \ \ \ \ Infection\ duration \ \ \ \ \ \ Transmission\ probability -->
<!-- $$ -->

```{r}
output$myMathText <- renderUI({
  val <- .75 * input$input_reco * input$input_trans
  withMathJax(HTML(paste0("$$R_{0} = ", val, "$$")))
})

uiOutput("myMathText")
```